---
title: "p8105_mtp_cz2634"
author: "Chenxin Zhang"
date: "10/20/2020"
output: github_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(gridExtra)
```


```{r}
dog_df = read_excel("./data/dogweights_07sept2020.xlsx")
```

# Probalem 1
**Introduction and Goal of the Report**

The raw data record the weights of three dogs named "Simon, Gagne, and Raisin" over 2 years with a standard object weighed on a dedicated scale. There are `r nrow(dog_df)` rows and `r ncol(dog_df)` columns. Observations are weight records of the three dogs through time.  It also includes occasional notes mainly  is time spent in a kennel.

The goal of the report is to 

**Import, tidy, and clean the data**
* create a tibble for month name change
* "read_excel" read the excel, "skip" skip the first raw, "col_types" set the column type  
* use function clean_names() in janitor packages to clean data  
* drop the row based on the missing values in var. 'raisin,simone, gagne'
* change the variable type by mutate()
* join two data frame by left_join()
* delete columns by select(-*)
```{r month_name}
month_df = tibble(month = 1:12, month_name = month.name) 
```

```{r year_month_date}
#separate the variable "date" to three columns "year, month, date"
dog_df = read_excel("./data/dogweights_07sept2020.xlsx", skip = 1, col_types = c("date", "guess", "guess", "guess", "guess", "text")) %>% 
  janitor::clean_names() %>% 
  drop_na(raisin) %>% 
  separate(date, c("year", "month", "date")) %>% 
  mutate(month = as.integer(month)) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month, -x6) %>% 
  select(year, month_name, everything()) 
```

* change the weights unit from lbs_oz to lbs for three dogs and the standard object
* separate one column to two by separate()
* transfer oz to lbs by oz/16

```{r}
weight_df = dog_df %>% 
  #change the weights unit from lbs_oz to labs for dog raisin
  separate(raisin, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    raisin_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(simone, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    simone_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(gagne, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    gagne_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(std, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    std_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
print()
```


**A dataframe include notes with date**
* change the variable name to "notes" by mutate()
* drop the row based on the missing values in var. 'notes' 

```{r}
notes_df = read_excel("./data/dogweights_07sept2020.xlsx", skip = 1, col_types = c("date", "guess", "guess", "guess", "guess", "text")) %>% 
  janitor::clean_names() %>% 
  drop_na(raisin) %>%
  mutate(notes = x6) %>%
  drop_na(notes) %>%  
  select(date, notes) %>% 
  separate(date, c("year", "month", "date")) %>% 
  mutate(month = as.integer(month)) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month) %>% 
  select(year, month_name, everything())
print(notes_df)
```

**export data frame**
* export the two data fram named weight_df and notes_df as CSVs by function write.csv. 
* save the CSVs to the same directory as the raw data in working directory by giving the path

```{r export df}
write.csv(weight_df, "./data/weight.xlsx")
write.csv(notes_df, "./data/notes.xlsx")
```

