---
title: "p8105_mtp_cz2634"
author: "Chenxin Zhang"
date: "10/20/2020"
output: github_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
 warning = FALSE, message = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")
library(tidyverse)
library(readxl)
library(gridExtra)
library(ggridges)
library(patchwork)

#devtools :: install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
library(wordcountaddin)

theme_set(theme_minimal()) 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r echo = FALSE}
dog_df = read_excel("./data/dogweights_07sept2020.xlsx")
```

## Problem 1 Tidy and clean data
**Introduction and Goal of the Report**

The raw data record the weights of three dogs named "Simon, Gagne, and Raisin" over 2 years with a standard object weighed. There are `r nrow(dog_df)` rows and `r ncol(dog_df)` columns. The goal of the report is to analyze the weight distribution of the three dogs over time.

**Import, tidy, and clean the data**

```{r}
#change the weights unit from lbs_oz to lbs for three dogs and the standard object
#read the excel, "skip" skip the first raw, "col_types" set the column type  
#use function clean_names() in janitor packages to clean data  
#drop the row based on the missing values in var. 'raisin,simone, gagne

dog_df = read_excel("./data/dogweights_07sept2020.xlsx", skip = 1, col_types = c("date", "guess", "guess", "guess", "guess", "text")) %>% 
  janitor::clean_names() %>% 
  drop_na(raisin) %>% 
  select(-x6) %>% 
  separate(raisin, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    raisin_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(simone, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    simone_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(gagne, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    gagne_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) %>% 
  separate(std, into = c("lbs", "oz"), sep = 2) %>% 
  mutate(
    lbs = as.numeric(lbs),
    oz = as.numeric(oz),
    std_lbs = lbs + oz/16) %>% 
  select(-lbs, -oz) 
```

```{r month_name}
month_df = tibble(month = 1:12, month_name = month.name) 
```

```{r weight_df}
#transfer oz to lbs by oz/16
#separate the variable "date" to three columns "year, month, date"
#change the variable type by mutate()
#join two data frame by left_join()
#delete columns by select(-*)
weight_df = dog_df %>% 
  separate(date, c("year", "month", "date")) %>% 
  mutate(month = as.integer(month)) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month) %>% 
  select(year, month_name, everything()) %>% 
print()
```  
* The data frame 'weight_df' is more tidy and readable. It delete rows where there is no record for weight of all three dogs, but keep rows when at least one weight is recorded. 

**A dataframe that includes notes with date**

```{r}
#change the variable name to "notes" by mutate()
#drop the row based on the missing values in var. 'notes'
notes_df = read_excel("./data/dogweights_07sept2020.xlsx", skip = 1, col_types = c("date", "guess", "guess", "guess", "guess", "text")) %>% 
  janitor::clean_names() %>% 
  drop_na(raisin) %>%
  mutate(notes = x6) %>%
  drop_na(notes) %>%  
  select(date, notes) %>% 
  separate(date, c("year", "month", "date")) %>% 
  mutate(month = as.integer(month)) %>% 
  left_join(month_df, by = "month") %>% 
  select(-month) %>% 
  select(year, month_name, everything())
print(notes_df)
```
* In this part, we produce a new data frame that include notes over time. There are totall `r nrow(notes_df)` notes which records things happen to dog like moving place, hair cut, death of Gagne. All three dogs had hair cut on 10/15/2018. Raisin and Simone had another hair cut on 5/23/2020, while Gagne hadn't because of the pass away. 


**export data frame**  
```{r export df, echo = TRUE}
#export the two data frame named ‘weight_df’ and ‘notes_df’ as CSVs
#Save the CSVs to the same directory as the raw data in working directory by giving the relative path.
write.csv(weight_df, "./data/weight.csv")
write.csv(notes_df, "./data/notes.csv")
```

## Problem 2 Summary of weight
**Briefly describe the resulting tidy dataset containing weights**  
* The weights data recording the weights of the three dogs and a standard object has `r nrow(weight_df)` rows and `r ncol(weight_df)` columns.  There are `r nrow(weight_df)` unique dates are included in the dateset. By observation the raw data, there is a miss recorded value of weight for raisin. The weight of raisin on 8/14/2019 is 7.9375 lbs while the records on near dates are all about 17.93 lbs. To analyze the data accurately, the analysis below will drop the miss recorded value.

**summary of the weight**

```{r}
#calculate the mean and sd of each and every column, ignoring missing values
dog_df2 = 
  dog_df %>% 
  select(-std_lbs) %>% 
  pivot_longer(
    raisin_lbs:gagne_lbs,
    names_to = "dog_name",
    values_to = "weight_lbs") %>% 
  filter(weight_lbs != "7.9375") %>% #drop the value of 7.9375
  drop_na()
```

```{r}
dog_df3 = dog_df2 %>% 
  group_by(dog_name) %>%
  summarize(mean_weight = mean(weight_lbs, na.rm = TRUE),
            sd_weight = sd(weight_lbs, na.rm = TRUE),
            observation = n()) %>% 
  knitr::kable(digits = 1) #creat a table and the digits and put one place to the decimal
print(dog_df3)
```
* The data frame 'dog_df3'  shows a summary about the number of total observation, average weight and the standard deviation for each dog. There are `r nrow(dog_df$gagne_lbs)` observations for gagne, `r (nrow(dog_df$raisin_lbs))-1` valid observations for raisin, and `r nrow(dog_df$simone_lbs)` observations for Simone. The mean weight of Raisin is 18.3, while it will be 18.2 if we do not drop the miss record value '7.9375'. The mean weights of Gagne and Simone are very close which are 10.3 and 10.4 respectively. Both weight values of Gagne and Raisin have a SD of 0.7, and the SD of weight value of Simone is 0.4, which reflects weight values of Gagne and Raisin are more spread from mean than that of Simone. Simone has a relatively stable weight than the other.

## Problem 3 Visualization
**two-panel plot**

```{r weight_dis}
weight_dis = dog_df2 %>%
  group_by(dog_name, weight_lbs) %>% 
  summarize(count = n()) %>% 
ggplot( aes(x = weight_lbs, y = dog_name, fill = count)) +
  geom_density_ridges(
    aes(point_color = dog_name, point_fill = count, point_shape = dog_name),
    alpha = .5, point_alpha = 0.9, jittered_points = TRUE
  ) +
  scale_point_color_hue(l = 40) +
  scale_discrete_manual(aesthetics = "point_shape", values = c(15, 16, 17)) +
  scale_x_continuous(
    breaks = seq(8,20,2))+
  theme(legend.position = ("none"),
    plot.title = element_text(lineheight = 4, color = "black", size = 10),
    axis.text.y = element_text(vjust = 0.5, hjust = 1, size = 8),
    axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 1, size = 10))+
  labs(title = "The distribution of weights for each dog",
    x = "weight_lbs",
    y = "count") +
  viridis::scale_color_viridis(discrete = TRUE)
```

```{r weight_time}
weight_time = dog_df2 %>% 
  ggplot(aes(x = date, y = weight_lbs)) + 
  geom_point(aes(color = dog_name), alpha = .4, size = 1) +
  geom_smooth(aes(color = dog_name), se = FALSE) +
  scale_y_continuous(breaks = c(8, 10, 12, 14,16,18,20,22),
                     labels = c("8", "10", "12", "14", "16", "18", "20", "22"))+
  theme(legend.position = ("bottom"),
        legend.title = element_blank(),
    plot.title = element_text(lineheight = 4, color = "black", size = 10),
    axis.text.y = element_text(vjust = 0.5, hjust = 1, size = 8),
    axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 1, size = 10))+
  labs(title = "Weight distrubution over time",
       x = "date",
       y = "weight_lbs") +
  viridis::scale_color_viridis(
    discrete = TRUE
)
```

```{r}
(weight_dis + weight_time)+ plot_layout(widths = 10, heights = 20)
```


* By visualizing, we can more clearly see the weight difference among the three dogs, and the weight distribution over time. Raisin is almost two times heavy than the other two dogs. The right pannel shows when time close to the death of Gagne, he lost weight over time from average 10 lbs to finally 8 lbs. Raisin and Simmon have very similar weight gain or lose trend over the two years. When one gain/lose weight, the other also gain/lose weight. Ao a summary, the fluctuation of weight of Raisin and Simmon are smooth over time.






